Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline

### 2014-06-26 paper.tex
Edits

\IEEEPARstart{T}{here} is considerable interest in the electrical modelling of electrodes. \cite{Cogan2008,Brown2008,Guo2012,Troy2006}
In \cite{Franks2005} a linearised model was presented, and in \cite{ScottSingle2013} a non-linear model
suitable for use with the SPICE circuit simulator was presented.
The SPICE model usefully characterises an electrode in a given electrolyte with a small number of parameters.
One major reason for the interest in the electrical impedance of electrodes concerns the design of electronics intended for integration into pacemakers, cochlear implants, spinal cord stimulators, etc.
The design of successful circuits depends upon a good understanding of external load impedance, while maximisation of battery life is linked to the use of electrodes whose impedance is well understood.

Having a compact model of an electrode in the appropriate electrolyte allows circuit designers to simulate their designs with valid loads.
{
    \color{blue}
For example, the authors of \cite{Ethier2011} are concerned with the impedance presented by their electrode loads, and calculate power efficiency of their circuits.
This value is affected by the load impedance assumed.
Nevertheless, it is clear that they do not have a good idea of, nor a good way to simulate with, the load truly presented by an electrode in vivo or in vitro.
}
Knowing the effect upon the model of changes in the electrode or electrolyte would allow designers to anticipate circumstances that will affect the load seen by their circuits.

@@

It has been suggested that the primary coupling mechanism between platinum electrodes and physiological saline is the reversible evolution of gas, probably oxygen, at the platinum surface, which dissolves without bubble formation.\cite{Greatbatch1969} Understanding the exact mechanism of displacement or Faradaic charge transfer at the electrode surface is beyond the scope of this work. Rather, we wish to characterise the electrode-electrolyte interface using a small number of parameters that can be fed into an electrical model.

@@

\section{Displacement Parameters / CPE}
\label{sect:cpe}
Displacement currents are responsible for capacitive behaviour at interfacial boundaries.
They are brought about by the redistribution of charge in response to applied fields, e.g., a charged electrode will attract or repel $Cl^{-}$ from its surface and reorientate polar molecules in the surrounding solution.\cite{Merrill05} This capacitive behaviour may also be the result of electrode polarisation in the form of reversible Faradaic reactions at the surface of the electrode.
Such reactions involving water and platinum, as identified in \cite{Horch2004,Mohtashami2011,Merrill05}, include:
\begin{eqnarray}
    Pt + H_{2}O &\Leftrightarrow& PtO + 2 H^{+} + 2 e^{-}\\
    PtO + H_{2}O &\Leftrightarrow& PtO_{2} + 2 H^{+} + 2e^{-}\\
    Pt + H^{+} & \Leftrightarrow & Pt-H\\
    Pt + H_{2}O + e^{-} &\Leftrightarrow& Pt-H+OH^{-}
\end{eqnarray}

@@

\section{Conclusion}
\label{sect:conclusion}
We have used the parameters of a compact electrical model of an implantable electrode as a way of objectively comparing scenarios.
We have measured how changing the concentration of PBS affects the parameters of this model.
We have drawn direct relationships with concentration and conductivity sufficient to predict interface characteristics at arbitrary dilutions of PBS.
We found that the magnitude of the CPE element moves much more slowly with concentration whereas the resistance moved almost linearly.
Using DC measurements we conclude that the concentration of PBS has no effect on the onset of Faradaic conduction for the range of concentrations measured. We hypothesise that this is because the double-layer sets the effective concentration of the species involved in the Faradaic reactions, in a localised volume surrounding each electrode.

In-vivo measurements of platinum electrodes in sheep spine show that no single concentration of PBS matches both the CPE and the resistive characteristics at once.
These measurements also reveal a considerable amount of reactance (see transimpedance measurements in vivo) that is captured neither by PBS nor the model presented when using a purely resistive spreading resistance.
Nevertheless, a single solution of PBS is to be used as a human phantom, a one-tenth concentration solution (by mass) is a good compromise.

@@

The final model, including 205 lines of resistive mesh, is represented by a SPICE netlist comprised of 348 lines. Simulation of this model, using the freely available ngSpice circuit simulator, at 1000 frequencies takes approximately 1 second on a modern CPU.

@@

%\bibitem{Weiland2000}
%    Weiland~James~D. and Anderson~David~J.,
%    ``Chronic Neural Stimulation with Thin-Film, Iridium Oxide Electrodes'',
%    {\em IEEE Transactions on Biomedical Circuits and Systems},
%    vol.~47 no.~7, July 2000, pp911--918

>>>>

\IEEEPARstart{T}{here} is considerable interest in the electrical modelling of electrodes.~\cite{Cogan2008,Brown2008,Guo2012,Troy2006}
In \cite{Franks2005} a linearised model was presented, and in \cite{ScottSingle2013} a non-linear model suitable for use with the SPICE circuit simulator was presented.
The SPICE model usefully characterises an electrode in a given electrolyte with a small number of parameters.
One major reason for the interest in the electrical impedance of electrodes concerns the design of electronics intended for integration into pacemakers, cochlear implants, spinal cord stimulators, etc.
The design of successful circuits depends upon a good understanding of external load impedance, while maximisation of battery life is linked to the use of electrodes whose impedance is well understood.

Having a compact model of an electrode in the appropriate electrolyte allows circuit designers to simulate their designs with valid loads.
For example, the authors of \cite{Ethier2011} are concerned with the impedance presented by their electrode loads, and calculate power efficiency of their circuits.
This value is affected by the load impedance assumed.
Nevertheless, it is clear that they do not have a good idea of, nor a good way to simulate with, the load truly presented by an electrode in vivo or in vitro.
Knowing the effect upon the model of changes in the electrode or electrolyte would allow designers to anticipate circumstances that will affect the load seen by their circuits.

@@

It has been suggested that the primary coupling mechanism between platinum electrodes and physiological saline is the reversible evolution of gas, probably oxygen, at the platinum surface, which dissolves without bubble formation.~\cite{Greatbatch1969}
Understanding the exact mechanism of displacement or Faradaic charge transfer at the electrode surface is beyond the scope of this work.
Rather, we wish to characterise the electrode-electrolyte interface using a small number of parameters that can be fed into an electrical model.

@@

\section{Displacement Parameters / CPE}
\label{sect:cpe}
Displacement currents are responsible for capacitive behaviour at interfacial boundaries.
They are brought about by the redistribution of charge in response to applied fields, e.g., a charged electrode will attract or repel $Cl^{-}$ from its surface and reorientate polar molecules in the surrounding solution. \cite{Merrill05}
This capacitive behaviour may also be the result of electrode polarisation in the form of reversible Faradaic reactions at the surface of the electrode.
Such reactions involving water and platinum, as identified in \cite{Horch2004,Mohtashami2011,Merrill05}, include:
\begin{eqnarray}
    Pt + H_{2}O &\Leftrightarrow& PtO + 2 H^{+} + 2 e^{-}\\
    PtO + H_{2}O &\Leftrightarrow& PtO_{2} + 2 H^{+} + 2e^{-}\\
    Pt + H^{+} & \Leftrightarrow & Pt-H\\
    Pt + H_{2}O + e^{-} &\Leftrightarrow& Pt-H+OH^{-}
\end{eqnarray}

@@

\section{Conclusion}
\label{sect:conclusion}
We have used the parameters of a compact electrical model of an implantable electrode as a way of objectively comparing scenarios.
We have measured how changing the concentration of PBS affects the parameters of this model.
We have drawn direct relationships with concentration and conductivity sufficient to predict interface characteristics at arbitrary dilutions of PBS.
We found that the magnitude of the CPE element moves much more slowly with concentration whereas the resistance moved almost linearly.
Using DC measurements we conclude that the concentration of PBS has no effect on the onset of Faradaic conduction for the range of concentrations measured.
We hypothesise that this is because the double-layer sets the concentration of species involved in the Faradaic reactions within a volume surrounding each electrode.

In-vivo measurements using platinum electrodes in sheep spine show that no single concentration of PBS matches both the CPE and the resistive characteristics at once.
These measurements also reveal a considerable amount of reactance (see transimpedance measurements in vivo) that is captured neither by PBS nor the model presented when using a purely resistive spreading resistance.
Nevertheless, if a single solution of PBS is to be used as a human phantom, a one-tenth concentration solution (by mass) is a good compromise.

@@

The final model, including 205 lines of resistive mesh, is represented by a SPICE netlist comprised of 348 lines.
A simulation of this model, using the freely available ngSpice circuit simulator, at 1000 frequencies takes approximately 1 second on a modern CPU.

@@
